### 2022-05-07 documentation/paper.tex
-

\documentclass{article}
\usepackage{neurips_2022} % for first submission 
% \usepackage[preprint]{neurips_2022}
% \usepackage[final]{neurips_2022}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
% custom
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}

\title{A Framework for Predictable Actor-Critic Control}
\author{%
    % Josiah Coad, \textsuperscript{\rm 1}
    % Jeff Hykin, \textsuperscript{\rm 2}
    % James Ault, \textsuperscript{\rm 2}
    % Guni Sharon \textsuperscript{\rm 2}
    Jeff Hykin
    \texttt{hippo@cs.cranberry-lemon.edu} \\
    % examples of more authors
    % \And
    % Coauthor \\
    % Affiliation \\
    % Address \\
    % \texttt{email} \\
    % \AND
    % Coauthor \\
    % Affiliation \\
    % Address \\
    % \texttt{email} \\
    % \And
    % Coauthor \\
    % Affiliation \\
    % Address \\
    % \texttt{email} \\
    % \And
    % Coauthor \\
    % Affiliation \\
    % Address \\
    % \texttt{email} \\
}

\begin{document}
    \maketitle
    \begin{abstract}
        FIXME
    \end{abstract}

    % 
    % epsilon optimizer
    % 
    \begin{algorithm}[h]
        % \SetAlgoLined\
        \SetKwInOut{Input}{Input}
        \SetKwInOut{Output}{Output}
        \Input{
            multiplier for epsilon $\mathrm{inc}$,\\
            baseline min confidence interval $\mathrm{b_{min}}$,\\
            confidence interval $\mathrm{ci_{lvl}}$,\\
            minimum average episode reward $\mathrm{b_{lower}}$,\\
            baseline confidence interval width $\mathrm{b_{ci}}$,\\
            number of baseline samples $\mathrm{b_{sz}}$,\\
            initial epsilon $\mathrm{\epsilon_{i}}$,\\
            initial horizon $\mathrm{h_{i}}$,\\
            number of iterations of refinement $\mathrm{iter}$\\    
        }
        \Output{
            \\
            a refined epsilon $\mathrm{\epsilon_{i}}$
            a refined horizon $\mathrm{h_{i}}$
        }
        \caption{Refine2}
        \label{algo:replan2}

        Initialize an empty array $ epsilon\_attempts $
        Initialize an empty hash map $ planlens_\epsilon $. Requesting a non-existent key for $ planlens_\epsilon $ returns an array containing $h_i$

        \For{$i$ in $[0,..., iter]$}{
            \label{ln:loopplan1}
            Initialize an empty array $sampled\_rewards$ 
            $a \gets 10$ 
            \For{$i$ in $[0,..., b_{sz}]$}{
                \label{ln:loopsample1}
                $discounted\_rewards, planlens 
                    \gets run\_episode(\epsilon=e_n)$ 
                $sampled\_rewards.push(\sigma(discounted\_rewards))$ 
                $planlens_\epsilon[\epsilon\_n] 
                    \gets concat(planlens_\epsilon[\epsilon_i], planlens)$

                \If{$length(sampled\_rewards) < 2$} {
                    $continue$
                }

                $s_{max}           \gets
                \hspace{0.5cm}    confidence\_interval\_upper(ci_{lvl}, sampled\_rewards)$
                $confidence\_size \gets
                \hspace{0.5cm}    confidence\_interval\_size(ci_{lvl} , sampled\_rewards)$
                \If{$s_{max} < baseline_{min} \leq confidence_size < b_ci$}{
                    $break$
                }
            }

            \If{$mean(sampled\_rewards) \geq b_{lower}$}{
                $\epsilon_n \*= inc$
            }{
                $\epsilon_n \/= inc$
            }

            $epsilon\_attempts.push(\epsilon_n)$
        }
        $\epsilon_o = median(epsilon\_attempts)$

        $h_o = max(median(planlens_\epsilon[\epsilon_o]), 1) * 2$

        \Return{
            $\mathrm{\epsilon_{o}}$
            $\mathrm{h_{o}}$
        }
    \end{algorithm}

    % TODO: refernce format: Citations may be author/year or numeric. Use natbib \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}

    % for graphics:
    %    \usepackage[pdftex]{graphicx} ...
    %    \includegraphics[width=0.8\linewidth]{myfile.pdf}
    % (\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})

    % 
    % 
    % 
    % Acknowledgments
    % 
    % 
    % 
    \begin{ack}
        FIXME
        Do {\bf not} include this section in the anonymized submission, only in the final paper.

        Use unnumbered first level headings for the acknowledgments.
        All acknowledgments go at the end of the paper before the list of references.
        - declare funding (financial activities supporting the submitted work)
        - competing interests (related financial activities outside the submitted work)
    \end{ack}

    % 
    % 
    % 
    % References
    % 
    % 
    % 
    \section*{References}
        Use unnumbered first-level heading for references.
        Any choice of citation style is acceptable as long as you are consistent.
        Reference section does not count towards the page limit.
    \medskip

    % 
    % 
    % 
    % Checklist
    % 
    % 
    % 
    \section*{Checklist}
    % change the default \answerTODO{} to \answerYes{}, \answerNo{}, or \answerNA{}
    % examples:
    % \begin{itemize}
    %   \item Did you include the license to the code and datasets? \answerYes{See Section~\ref{gen_inst}.}
    %   \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
    %   \item Did you include the license to the code and datasets? \answerNA{}
    % \end{itemize}
    \begin{enumerate}

        \item For all authors...
            \begin{enumerate}
                \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
                    \answerTODO{}
                \item Did you describe the limitations of your work?
                    \answerTODO{}
                \item Did you discuss any potential negative societal impacts of your work?
                    \answerTODO{}
                \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
                    \answerTODO{}
            \end{enumerate}

        \item If you are including theoretical results...
            \begin{enumerate}
                \item Did you state the full set of assumptions of all theoretical results?
                    \answerTODO{}
                \item Did you include complete proofs of all theoretical results?
                    \answerTODO{}
            \end{enumerate}

        \item If you ran experiments...
            \begin{enumerate}
                    \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
                        \answerTODO{}
                    \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
                        \answerTODO{}
                    \item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
                        \answerTODO{}
                    \item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
                        \answerTODO{}
            \end{enumerate}

        \item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
            \begin{enumerate}
                \item If your work uses existing assets, did you cite the creators?
                    \answerTODO{}
                \item Did you mention the license of the assets?
                    \answerTODO{}
                \item Did you include any new assets either in the supplemental material or as a URL?
                    \answerTODO{}
                \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
                    \answerTODO{}
                \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
                    \answerTODO{}
            \end{enumerate}

        \item If you used crowdsourcing or conducted research with human subjects...
            \begin{enumerate}
                \item Did you include the full text of instructions given to participants and screenshots, if applicable?
                    \answerTODO{}
                \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
                    \answerTODO{}
                \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
                    \answerTODO{}
            \end{enumerate}

    \end{enumerate}

    % 
    % 
    % Appendix
    % 
    % 
    \appendix
    \section{Appendix}
    Optionally include extra information (complete proofs, additional experiments and plots) in the appendix.
    This section will often be part of the supplemental material.

\end{document}