### 2022-05-09 documentation/paper.tex
improve math stuff

\documentclass{article}
\usepackage{neurips_2022} % for first submission 
% \usepackage[preprint]{neurips_2022}
% \usepackage[final]{neurips_2022}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
% custom
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}

\title{A Framework for Predictable Actor-Critic Control}
\author{%
    % Josiah Coad, \textsuperscript{\rm 1}
    % Jeff Hykin, \textsuperscript{\rm 2}
    % James Ault, \textsuperscript{\rm 2}
    % Guni Sharon \textsuperscript{\rm 2}
    Jeff Hykin
    \texttt{hippo@cs.cranberry-lemon.edu} \\
    % examples of more authors
    % \And
    % Coauthor \\
    % Affiliation \\
    % Address \\
    % \texttt{email} \\
    % \AND
    % Coauthor \\
    % Affiliation \\
    % Address \\
    % \texttt{email} \\
    % \And
    % Coauthor \\
    % Affiliation \\
    % Address \\
    % \texttt{email} \\
    % \And
    % Coauthor \\
    % Affiliation \\
    % Address \\
    % \texttt{email} \\
}

\begin{document}
    \maketitle
    \begin{abstract}
        FIXME
    \end{abstract}

    % 
    % epsilon optimizer
    % 
    \begin{algorithm}[h]

        \newcommand{\planlensPerEpsilon}{planlens_\epsilon}
        \newcommand{\discountedRewards}{discounted\_rewards}
        \newcommand{\sampledRewards}{sampled\_rewards}
        \newcommand{\sum}{\sigma}
        \newcommand{\minimumAcceptableReward}{b_{lower}}
        \newcommand{\numberOfbaselineSamples}{b_{sz}}
        \newcommand{\initialEpsilon}{\epsilon_{i}}
        \newcommand{\runningEpsilon}{\epsilon_{r}}
        \newcommand{\finalEpsilon}{\epsilon_{o}}
        \newcommand{\allEpsilons}{a_\epsilon}
        \newcommand{\intialHorizon}{\h_{i}}
        \newcommand{\finalHorizon}{\h_{o}}
        \newcommand{\planLengths}{planlens}

        % \SetAlgoLined\
        \SetKwInOut{Input}{Input}
        \SetKwInOut{Output}{Output}
        \Input{
            $\mathrm{\minimumAcceptableReward}$, minimum acceptable reward \\
            $\mathrm{iter}$ , number of iterations of refinement \\
            $\mathrm{inc}$, scale multiplier for epsilon \\
            $\mathrm{\numberOfbaselineSamples}$, number of baseline samples \\
            $\mathrm{b_{min}}$, min confidence interval of baseline \\
            $\mathrm{b_{ci}}$, baseline confidence interval width \\
            $\mathrm{ci_{lvl}}$, confidence level  \\
            $\mathrm{\initialEpsilon}$, initial epsilon \\
            $\mathrm{\intialHorizon}$, initial horizon \\
        }
        \Output{
            \\
            $\mathrm{\finalEpsilon}$, a refined epsilon\\
            $\mathrm{\finalHorizon}$, a refined horizon\\
        }
        \caption{Find Epsilon}
        \label{algo:find_epsilon}
        \vspace{8pt}

        \(\runningEpsilon \gets  \initialEpsilon \) \\
        \(\allEpsilons \gets []\) \# empty array \\
        \(\planlensPerEpsilon \gets \{\}\) \# empty hashmap \\
        \# Note: requesting a non-existent key for this hash-map returns an array containing $\intialHorizon$ \\
        \For{$i$ in $[0,..., iter]$}{
            % \label{ln:loopplan1}
            \(\sampledRewards \gets []\) \# empty array \\
            \For{$j$ in $[0,..., \numberOfbaselineSamples]$}{
                % \label{ln:loopsample1}
                $\discountedRewards, \planLengths \gets runEpisode(\epsilon=\runningEpsilon)$ 
                $\sampledRewards.push(\sum(\discountedRewards))$ 
                $\planlensPerEpsilon[\runningEpsilon] \gets concat(\planlensPerEpsilon[\runningEpsilon], \planLengths)$ 
                \If{$length(\sampledRewards) < 2$} {
                    $continue$
                }

                $s_{max}, s_{min} \gets confidenceInterval(ci_{lvl}, \sampledRewards)$
                $s_{ci} \gets s_{max} - s_{min}$
                \If{$s_{max} < \minimumAcceptableReward \texttt{or} s_{ci} < b_ci$}{
                    $break$
                }
            }

            \If{$mean(\sampledRewards) \geq \minimumAcceptableReward$}{
                $\runningEpsilon \gets \runningEpsilon \* inc$
            }{
                $\runningEpsilon \gets \runningEpsilon / inc$
            }

            $\allEpsilons.push(\runningEpsilon)$
        }
        $\finalEpsilon \gets median(\allEpsilons)$

        $\finalHorizon \gets max(median(\planlensPerEpsilon[\finalEpsilon]), 1) * 2$

        \Return{
            $\mathrm{\finalEpsilon}$
            $\mathrm{\finalHorizon}$
        }
    \end{algorithm}

    % TODO: refernce format: Citations may be author/year or numeric. Use natbib \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}

    % for graphics:
    %    \usepackage[pdftex]{graphicx} ...
    %    \includegraphics[width=0.8\linewidth]{myfile.pdf}
    % (\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})

    % 
    % 
    % 
    % Acknowledgments
    % 
    % 
    % 
        % \begin{ack}
        %     FIXME
        %     Do {\bf not} include this section in the anonymized submission, only in the final paper.

        %     Use unnumbered first level headings for the acknowledgments.
        %     All acknowledgments go at the end of the paper before the list of references.
        %     - declare funding (financial activities supporting the submitted work)
        %     - competing interests (related financial activities outside the submitted work)
        % \end{ack}

    % 
    % 
    % 
    % References
    % 
    % 
    % 
    \section*{References}
        Use unnumbered first-level heading for references.
        Any choice of citation style is acceptable as long as you are consistent.
        Reference section does not count towards the page limit.
    \medskip

    % 
    % 
    % 
    % Checklist
    % 
    % 
    % 
    \section*{Checklist}
    % change the default \answerTODO{} to \answerYes{}, \answerNo{}, or \answerNA{}
    % examples:
    % \begin{itemize}
    %   \item Did you include the license to the code and datasets? \answerYes{See Section~\ref{gen_inst}.}
    %   \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
    %   \item Did you include the license to the code and datasets? \answerNA{}
    % \end{itemize}
    \begin{enumerate}

        \item For all authors...
            \begin{enumerate}
                \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
                    \answerTODO{}
                \item Did you describe the limitations of your work?
                    \answerTODO{}
                \item Did you discuss any potential negative societal impacts of your work?
                    \answerTODO{}
                \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
                    \answerTODO{}
            \end{enumerate}

        \item If you are including theoretical results...
            \begin{enumerate}
                \item Did you state the full set of assumptions of all theoretical results?
                    \answerTODO{}
                \item Did you include complete proofs of all theoretical results?
                    \answerTODO{}
            \end{enumerate}

        \item If you ran experiments...
            \begin{enumerate}
                    \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
                        \answerTODO{}
                    \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
                        \answerTODO{}
                    \item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
                        \answerTODO{}
                    \item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
                        \answerTODO{}
            \end{enumerate}

        \item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
            \begin{enumerate}
                \item If your work uses existing assets, did you cite the creators?
                    \answerTODO{}
                \item Did you mention the license of the assets?
                    \answerTODO{}
                \item Did you include any new assets either in the supplemental material or as a URL?
                    \answerTODO{}
                \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
                    \answerTODO{}
                \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
                    \answerTODO{}
            \end{enumerate}

        \item If you used crowdsourcing or conducted research with human subjects...
            \begin{enumerate}
                \item Did you include the full text of instructions given to participants and screenshots, if applicable?
                    \answerTODO{}
                \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
                    \answerTODO{}
                \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
                    \answerTODO{}
            \end{enumerate}

    \end{enumerate}

    % 
    % 
    % Appendix
    % 
    % 
    \appendix
    \section{Appendix}
    Optionally include extra information (complete proofs, additional experiments and plots) in the appendix.
    This section will often be part of the supplemental material.

\end{document}